---
title: "An analysis on the influences between programming languages"
author: Eros Fabrici
date: 05/02/21
output:
    ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =FALSE, cache = TRUE, message = FALSE,
                      warning = FALSE, fig.align = 'center', out.width = 1000, out.height = 550, fig.width = 10, fig.height = 8)
```
<!-- Script for adding ToC !-->
<script>
document.addEventListener('DOMContentLoaded', function() {
  TableOfContents();
}
);

function TableOfContents(container, output) {

var output = output || '#toc';

// Get all elements with class: section or subsection
idfromclass = document.querySelectorAll('.section');

    // Create the list element:
    var list = document.createElement('ul');

    // Iterate through all found elements
    for(var i = 0; i < idfromclass.length; i++) {

        // Create the list item:
        var item = document.createElement('li');

        // Set its contents:
        var id = idfromclass[i].id

        // Replace - in id with whitespace
        var titleText = id.replace(/-/gi, " ");

        // Add text to list element
        item.appendChild(document.createTextNode(titleText));

        // Add it to the list:
        list.appendChild(item);
    }

// Return generated HTML to toc div in slide
document.querySelector(output).innerHTML = list.innerHTML;

// Generate instruction message if no classes are defined
if (idfromclass.length == 0) { document.querySelector(output).innerHTML = "Add {.section} or {.subsection} to slide name to generate TOC"; }

};
</script>

```{r libraries}
source('clean_data.R')
library(tidyverse)
library(igraph)
library(ggraph)
library(devtools)
install_github('eferos93/tidygraph')
library(tidygraph)
library(tweenr)
set_graph_style()
graph <- tbl_graph(nodes = programming_languages_tibble,
                   edges = influences_tibble, directed = TRUE) %>%
  activate(nodes) %>%
  arrange(year)
```

## Table of content

<div id="toc"></div>


# Introduction {.section}
## Dataset
* Dataset can be found in the [ICON](https://icon.colorado.edu/#!/) website
* The dataset represents a directed graph, and it describes the influences between the programming languages released between 1952 and 2010
* It has 395 nodes and 762 edges

---
```{r, eval = FALSE}
# years <- graph %>%
#  as_tibble(what = 'vertices') %>%
#   select(year) %>%
#   filter(!is.na(year)) %>%
#   distinct() %>%
#   arrange(year)
# # year_tween <- tween_states(years, tweenlength = 1,
# #                            statelength = 1, ease = 'cubic-in-out',
# #                            nframes = length(years) * 16 + 8
# # )
# # title_transp <- tween_t(c(0, 1, 0, 0, 0), 16, 'cubic-in-out')[[1]]
# anim_graph <- graph %>% activate(nodes) %>%  mutate(ease = 'linear')
# year_tween <- tween_elements(anim_graph, time = 'year', group = 'ease',
#                              ease = 'ease', nframes = 59 * 20)
# graph_tween_appear <- tween_appear(year_tween, time = 'year', nframes = 59 * 20) %>% keep_state(20)
#
# plot_data <- graph_tween_appear %>% filter(.frame = i)
# p <- ggraph(plot_data, layout = 'fr') +
#   geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE, edge_colour = 'grey66') +
#   geom_node_point(show.legend = FALSE) +
#   geom_node_text(aes(label = name), repel = TRUE, show.legend = FALSE,
#                  colour = 'red', family = 'serif') +
#   theme_graph()
# plot(p)
#
# # for (i in seq_len(length(years) * 16)) {
# #   years_to_plot <- 1952:year_tween[year_tween$.frame == i,]
# #   filtered_graph <-
# #     graph %>%
# #       activate(nodes) %>%
# #       filter(year %in% years_to_plot)
# #   set.seed(10)
# #   p <- ggraph(filtered_graph, layout = 'fr') +
# #     geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE, edge_colour = 'grey66') +
# #     geom_node_point(show.legend = FALSE) +
# #     geom_node_text(aes(label = name), repel = TRUE, show.legend = FALSE, colour = 'red', family = 'serif') +
# #     theme_graph()
# #   plot(p)
# # }
```
```{r The Network}
graph %>% filter(!node_is_isolated()) %>%
  ggraph(layout = 'fr') +
  geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE, edge_colour = 'grey66') +
  geom_node_point(show.legend = FALSE) +
  geom_node_text(aes(label = name), repel = TRUE, show.legend = FALSE, colour = 'red', family = 'serif')
```

## Questions
* The aim of this project is to answer the following questions by analysing the network:
    + What are the most influential programming languages?
    + Which are the most influenced?
    + Does the programming languages of the same paradigm family influence and get influenced between each other the most?
    + Are the programming languages of the same paradigm family more influenced between each other rather than the others belonging to other paradigms?

# Network Analysis {.section}
## Most influential programming languages

# Conclusions {.section}

## References {.section}

```{r Centrality helper functions, eval = FALSE}
get_centrality <- function (graph, centrality_function) {
  result_centraility <-
          graph %>%
          activate(nodes) %>%
          mutate(influence = centrality_function)
  return(result_centraility)
}

plot_centrality <- function (graph_with_centrality, filtering_degree = 0) {
  graph_with_centrality %>%
          filter(influence > filtering_degree) %>%
          filter(!node_is_isolated()) %>%
          ggraph(layout = 'fr') +
          geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE, edge_colour = 'grey66') +
          geom_node_point(aes(size = influence), show.legend = FALSE) +
          geom_node_text(aes(label = name), repel = TRUE, show.legend = FALSE, colour = 'red', family = 'serif') +
          theme_void()
}

plot_most_influent_languages <- function (graph_with_centraility, how_many = 10, centrality_metric) {
  as_tibble(graph_with_centraility, what = 'vertices') %>%
          top_n(how_many) %>%
          select(name, influence) %>%
          ggplot(aes(x = reorder(as.factor(name), -influence), y = influence)) +
          geom_bar(stat = 'identity', fill = 'blue4') +
          labs(
                  title = 'Top ten most influential programming languages',
                  subtitle = paste('Centrality metric used:', centrality_metric),
                  x = '',
                  y = 'Influence'
          ) +
          theme_classic()
}

subgraph_of_programming_language <- function (graph, programming_language,
                                              selected_mode = 'all') {
  subgraph <- graph %>%
          convert(to_bfs_tree,
                  which(.N()$name == programming_language), mode = selected_mode) %>%
          activate(nodes) %>%
          filter(!node_is_isolated())

  subgraph %>%
          ggraph(layout = 'fr') +
          geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE, edge_colour = 'grey66') +
          geom_node_point(aes(size = influence), show.legend = FALSE) +
          geom_node_text(aes(filter = name != programming_language, label = name),
                         repel = TRUE, show.legend = FALSE, colour = 'red', family = 'serif') +
          geom_node_text(aes(filter = name == programming_language, label = name),
                         colour = 'green', family = 'serif', repel = TRUE, show.legend = FALSE) +
          theme_void()
}
```

```{r Centrality Degree In, eval = FALSE}
graph_influence_degree <- get_centrality(graph, centrality_degree(mode = 'in'))
plot_centrality(graph_influence_degree)
plot_most_influent_languages(graph_influence_degree, centrality_metric = 'in-degree')
```

```{r Centrality Degree Out, eval = FALSE}
graph_influence_degree <- get_centrality(graph, centrality_degree(mode = 'out'))
plot_centrality(graph_influence_degree)
plot_most_influent_languages(graph_influence_degree, centrality_metric = 'out-degree')
```
```{r Centrality Closeness, eval = FALSE}
graph_closeness_centrality <- get_centrality(graph, centrality_closeness(mode = 'in'))
plot_centrality(graph_closeness_centrality)
plot_most_influent_languages(graph_closeness_centrality, centrality_metric = 'closeness')
```
```{r Centrality Betweeness, eval = FALSE}
graph_edge_betweenness <- get_centrality(graph, centrality_betweenness())
plot_centrality(graph_edge_betweenness)
plot_most_influent_languages(graph_edge_betweenness, centrality_metric = 'betweenness')
```
```{r Subgraphs, eval = FALSE}
subgraph_of_programming_language(graph_edge_betweenness, 'python', 'out')
subgraph_of_programming_language(graph_edge_betweenness, 'ruby')
subgraph_of_programming_language(graph_edge_betweenness, 'haskell', 'out')
subgraph_of_programming_language(graph_edge_betweenness, 'c')
```
```{r subgraphs with betweenness, eval = FALSE}
selected_mode <- 'out'
subgraph_of_programming_language(graph_edge_betweenness, 'python', selected_mode)
subgraph_of_programming_language(graph_edge_betweenness, 'ruby', selected_mode)
subgraph_of_programming_language(graph_edge_betweenness, 'haskell', selected_mode)
subgraph_of_programming_language(graph_edge_betweenness, 'c', selected_mode)
```
```{r Centrality Authority, eval = FALSE}
graph_authority <- get_centrality(graph, centrality_authority())
plot_centrality(graph_authority)
plot_most_influent_languages(graph_authority, centrality_metric = 'authority')
```
```{r Centrality Hub, eval = FALSE}
graph_hub <- get_centrality(graph, centrality_hub())
plot_centrality(graph_hub)
plot_most_influent_languages(graph_hub, centrality_metric = 'hub')
```
```{r Centrality Pagerank, eval = FALSE}
graph_pagerank <- get_centrality(graph, centrality_pagerank())
plot_centrality(graph_pagerank)
plot_most_influent_languages(graph_pagerank, centrality_metric = 'pagerank')
```
# Clustering
I linguaggi dello stesso paradigma si influenzano pricipalmente tra di loro?

```{r Clustering helper functions}
get_clusters <- function (graph, centrality_function, clustering_function) {
  graph_no_isolated_nodes <- graph %>%
          activate(nodes) %>%
          mutate(influence = centrality_function) %>%
          filter(!node_is_isolated()) %>%
          filter(influence > 0)

  clustered <- graph_no_isolated_nodes %>%
          activate(nodes) %>%
          mutate(group = clustering_function %>%  as.factor())
  return(clustered)
}
plot_clusters <-  function(clustered, centrality_function_name, clustering_fuction_name, groups = NULL) {
    if(!is.null(groups)) {
      clustered <- clustered %>% filter(group %in% groups)
    }
    clustered %>%
          ggraph(layout = 'fr') +
          geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE, edge_colour = 'grey66') +
          geom_node_point(aes(colour = group, size = influence), show.legend = FALSE) +
          geom_node_text(aes(label = name), repel = TRUE, show.legend = FALSE, colour = 'black', family = 'serif') +
          facet_nodes(~group) +
          labs(
                  title = 'Community detection',
                  subtitle = paste0('Centrality metric used to filter nodes: ', centrality_function_name,
                                    '. Clustering algorithm used: ', clustering_fuction_name)
          ) +
          theme_graph(foreground = 'steelblue', fg_text_colour = 'white')
}
```
```{r Clustering 1, eval = FALSE}
get_clusters(graph, centrality_pagerank(), group_infomap()) %>% plot_clusters('pagerank', 'infomap', 1:8)
get_clusters(graph, centrality_betweenness(), group_infomap()) %>% plot_clusters('betweenness', 'infomap', 1:8)
```
```{r Clustering 2, eval = FALSE}
get_clusters(graph, centrality_closeness(), group_infomap()) %>% plot_clusters('closeness', 'infomap', 1:4)
get_clusters(graph, centrality_betweenness(), group_infomap()) %>% plot_clusters('betweenness', 'infomap', 1:4)
```